---
title: "Bus Stop Finding"
output: html_document
date: "2023-11-16"
---

```{r setup, include=FALSE}

library('GISTools')
library('raster')
library('tmap')
library('rgdal') 
library('maptools') 
library('raster') 
library(raster)
library(sp)
library(sf)
library(tidyverse)
library(rgeos)
library(ggplot2)
library(psych)
library(spatstat)
library(spdep)
library(jsonlite)
library(gstat)
library(magick)


```


# Finding optimal bus stops

## Creating base layer map

Here we also set the crs to WGS84 as it causes calculation later to be simple as other data is left as lat lon coordinated.


```{r base}

#If we do not want the subzone lines
subzones <- st_read(dsn = "dataset/basemap", layer = "MP14_SUBZONE_NO_SEA_PL")
base_map_no_border <- st_read(dsn = "dataset/basemap", layer = "SGP_adm0")
base_map_no_border <- base_map_no_border %>% select(ISO)
base_map_no_border <- st_transform(base_map_no_border, crs=4326)

#If we want the subzone lines
base_map <- st_read(dsn = "dataset/basemap", layer = "MP14_SUBZONE_NO_SEA_PL")
base_map <- st_transform(base_map, crs=4326)
base_map <- st_make_valid(base_map)

```


## Loading data

In this file, we will be loading 4 main datasets:
- Youth density
- HDB locations
- Bus Stop locations
- Nightlife Locations


```{r load}

#Youth density
yth_data <- read.csv("dataset/filteredSG_population_density.csv") %>% select(longitude, latitude, youth)
yth_spatial <- st_as_sf(yth_data, coords = c("longitude", "latitude"), crs = 4326)


#HDB Locations
hdb <- read.csv("dataset/hdb.csv") %>% mutate(addr = paste(blk_no,street)) %>% select(addr,lat,lng)
hdb_spatial <- st_as_sf(hdb, coords = c("lng", "lat"), crs = 4326)


#Read & Load Bus Stop Data
bus_data <- data.frame()
for(i in 0:10) {
  file_name <- paste0("dataset/busstops/response(", i, ").json")
  # Handle the case for response.json (without the index)
  if(i == 0) {
    file_name <- "dataset/busstops/response.json"
  }
  json_read <- fromJSON(file_name, flatten = TRUE)
  bus_data <- rbind(bus_data, json_read$value)
}
bus_spatial <- st_as_sf(bus_data, coords = c("Longitude", "Latitude"), crs = 4326)


#Nightlife DataFrame
nightlife <- st_read('dataset/processed_datasets/nightlife_locations.geojson', crs = st_crs(subzones))
nightlife <- st_transform(nightlife, crs = st_crs(4326))

```


# HDBs Bus Stop Analysis

## Visualizing Youth Density

So the first step is to see how youth density looks like. Now what we notice is that youth density is a point plot, so let's see how that looks. 


```{r yth pts}
#Plot the points to visualise how the values look #plot takes a while to load
yth_spatial_map <- tm_shape(base_map_no_border) + tm_borders() + 
  tm_shape(yth_spatial) + tm_dots(size = 0.05, col = 'youth', title = 'Youth Density(%)')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Youth Densities(%) \nin Singapore",
            title.position = c('left', 'top'))
yth_spatial_map

tmap_save(yth_spatial_map, filename = "plots/optimal_stops/yth_spatial_map.png")

```

No what we realise is that the youth points are clustered around a certain value, so we can actually represent them as polygons instead. 

```{r yth polys}

#We notice that a lot of the same values are congregated together, so lets make them into polygon
#Group points with the same value and create convex hulls
grouped_points <- yth_spatial %>%
  group_by(youth) %>%
  summarize()
yth_poly <- grouped_points %>%
  st_convex_hull()
yth_poly <- yth_poly[-14,] #remove the individual point which is out of place 

#Lets see how this looks like
yth_poly_plot <- tm_shape(base_map_no_border) + tm_borders() + 
  tm_shape(yth_poly) + tm_polygons(col = 'youth', title = 'Youth Density') + 
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="HDB locations \nwithin Singapore",
            title.position = c('left', 'top'))
yth_poly_plot

tmap_save(yth_poly_plot, filename = "plots/optimal_stops/yth_poly_map.png")
```

Computation is MUCH faster now as we have reduced from 217054 data points to 32 polygons, great!  
Now lets visualise the HDBs. 

```{r hdb_yth poly}

joined <- st_join(hdb_spatial,yth_poly) #combine the HDB data with the youth density to assign youth density values to each HDB

#HDB point graph
hdb_point_plot <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_spatial) + tm_dots(size = 0.1, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "bottom"),
            legend.text.size = 0.75, 
            legend.title.size = 1.5,
            title="HDB locations \nwithin Singapore",
            title.position = c('left', 'top'))

hdb_point_plot

tmap_save(hdb_point_plot, filename = "plots/optimal_stops/hdb_point_plot.png")

```

Lastly, lets see the combination of both youth density and HDBs

```{r yth and hdbs}

hdb_yth_plot <- tm_shape(base_map) + tm_borders() + 
  tm_shape(yth_poly) + tm_polygons(col = 'youth', title = 'Youth Density') + 
    tm_shape(hdb_spatial) + tm_dots(size = 0.1, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="HDB locations \nwithin Singapore",
            title.position = c('left', 'top'))

hdb_yth_plot

tmap_save(hdb_yth_plot, filename = "plots/optimal_stops/hdb_yth_plot.png")
```

## Fixing data

So now our objective is to assing a youth density value to each HDB. However, what we noticed is that some HDBs (especially in the punggol area) are missing values and we also have some locations with multiple values. Now in order to fill in the data, rather than imputing with the mean, we decided that we can interpolate the data.  

In terms of deciding which interpolation method to use, we decided to use IDW as the appropriate method to fill in the data. This is becaus we wanted a deterministic model in order to fill in the missing data **consistently**. Furthermore, kringing would not perform well for this form of missing data which only has data from one side but not the other.  

So our first step is to create a simple dataframe which contains both the youth density data and HDB locations.

```{r joined}
joined <- st_join(hdb_spatial,yth_poly) #combine the HDB data with the youth density to assign youth density values to each HDB

#First, we need to seperate out data with NA values and only interpolate on coordinates with points

joined_noNA <- joined %>% filter(!is.na(youth))
joined_spatial <- as(joined_noNA, "Spatial")

```

We then create our raster grid for the IDW process.

```{r sg raster}

base_spatial <- as(base_map, 'Spatial')

#Creating a raster layer for interpolation
grd              <- as.data.frame(spsample(base_spatial, "regular", n=10000))
names(grd)       <- c("lat", "lng")
coordinates(grd) <- c("lat", "lng")
gridded(grd)     <- TRUE
fullgrid(grd)    <- TRUE

crs(grd) <- crs(base_spatial)

```

Now onto creating the IDW, determining a good idp value and plotting it out.

```{r idw}

yth.idw <- gstat::idw(youth ~ 1, joined_spatial, newdata=grd, idp=3.0) #idp of 3 chosen as we want closer values to mean more

#Find out what is a good idp value to use #10min runtime
IDW.out <- vector(length = length(joined_spatial))
for (i in 1:length(joined_spatial)) {
  IDW.out[i] <- gstat::idw(youth ~ 1, joined_spatial[-i,], joined_spatial[i,], idp=3.0)$var1.pred
}


#Plot the differences
OP <- par(pty="s", mar=c(4,3,0,0))
plot(IDW.out ~ joined_spatial$youth, asp=1, xlab="Observed", ylab="Predicted", pch=16,
     col=rgb(0,0,0,0.5))
abline(lm(IDW.out ~ joined_spatial$youth), col="red", lw=2,lty=2)
abline(0,1)
par(OP)

#Looks like the IDW looks valid

r_hdb_idw       <- raster(yth.idw)
r_hdb_idw.m     <- mask(r_hdb_idw, base_spatial) 


idw_full <- tm_shape(r_hdb_idw.m) + 
  tm_raster(n=10,palette = "Blues", stretch.palette = FALSE,title="Predicted Youth Density (%)") +
  tm_shape(base_map) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(joined_spatial) + tm_dots(size=0.005) +
  tm_legend(legend.outside=TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="IDW of HDB locations",
            title.position = c('left', 'top'))

idw_full

tmap_save(idw_full, filename = "plots/optimal_stops/idw_full.png")
```

## Filtering

Now that we have assigned each HDB a youth density value, the next step would be to filter out HDBs with lower youth density values.  

The first step to doing that is by visualising the IDW values in a histogram. We hace also calculated some key stats for the youth density values.

```{r hist}
#EDA on youth density, making a histogram of youth densities
yth_hist <- r_hdb_idw@data@values

hist <- ggplot(data.frame(values = yth_hist), aes(x = values)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black", aes(y = ..count..)) +
  ggtitle("Histogram of Youth Densities (%)") +
  xlab("Youth Density (%)") +
  ylab("Frequency") + theme_minimal()

hist

ggsave("plots/optimal_stops/hist.png",hist)

#A couple of statistics for youth density
mean(yth_hist)
median(yth_hist)
quantile(yth_hist,0.25)
quantile(yth_hist,0.75)
max(yth_hist)

```


The next step would be to determine a good threshold for youth density to filter out the HDB locations. Ideally, we would want to split the HDBs into 2 regions for the ideal bus stops to run well.  

Over a number of iterations, we set this filter to take the HDBs in the top 50% of youth densitys.  

Unfortunately, because assigning each HDB to the value from the raster grid is computationally intensive, we decided to reclassify the raster layer into 25 layers from 0 to 12.5 with a 0.5 interval. This can be seen in the chunk below.

```{r assigning yth}

#Before we move on, we realise that with spatial interpolation, the assigning of values will take a really long time
#As such, we reclassify the values into groups before converting them nto polygons for analysis

reclass_value_idw <- c(0,0.25,0) #to reclass values into 25 different bins
for (i in seq(0,24,1)) {
  reclass_value_idw[3*i + 4] = 0.25 + 0.5*i
  reclass_value_idw[3*i + 5] = 0.75 + 0.5*i
  reclass_value_idw[3*i + 6] = 0.5 + 0.5*i
}

reclass_r_hdb_idw <- reclassify(as(r_hdb_idw, "RasterLayer"), reclass_value_idw) #reclassify kde values to groups
filtered_r_hdb_idw <- reclass_r_hdb_idw
filtered_r_hdb_idw[filtered_r_hdb_idw <= quantile(yth_hist,0.5)] <- NA

#overlaid IDW polygon plot with hdb plot  
filtered_r_hdb_idw.m     <- mask(filtered_r_hdb_idw, base_spatial) 


idw_filter <- tm_shape(filtered_r_hdb_idw.m) + 
  tm_raster(n=10,palette = "Blues", stretch.palette = FALSE,title="Predicted Youth Density (%)") +
  tm_shape(base_map) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(joined_spatial) + tm_dots(size=0.005) +
  tm_legend(legend.outside=TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered IDW of  HDB locations",
            title.position = c('left', 'top'))

idw_filter

tmap_save(idw_filter, filename = "plots/optimal_stops/idw_filter.png")

```


Now what we are left is to remove the HDBs which lie outside the new raster layer. In order to accomplish this, we convert the raster values into polygons of equal value. Then an intersection is taken to fitler out the remaining HDBs. The resultant plot is as such.


```{r hdb filter}

#Converting raster layer into polygons to view overlay over Singapore
threshold <- quantile(yth_hist,0.5) #define a threshold to determine what values of youth density we drop
hdb_idw <- rasterToPolygons(reclass_r_hdb_idw, fun = function(x) {x > threshold},dissolve = TRUE) #
hdb_idw <- st_as_sf(hdb_idw)
hdb_idw <- hdb_idw %>% rename(youth = var1.pred)
hdb_idw <- st_cast(hdb_idw,'MULTIPOLYGON')
st_crs(hdb_idw) <- st_crs("EPSG:4326") #set crs for new polygons
hdb_idw <- st_intersection(hdb_idw, base_map) #to ensure new polygons stay within SG boundary

#assigning each hdb a youth value by intersecting the multipolygons to with all hdbs
hdb_yth_pts <- st_intersection(hdb_spatial, hdb_idw)  #takes some time

idw_final <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_idw) + tm_polygons(col = 'youth', title = 'Youth Density', palette = 'Blues') + 
    tm_shape(hdb_yth_pts) + tm_dots(size = 0.1, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered HDB locations",
            title.position = c('left', 'top'))

idw_final

tmap_save(idw_final, filename = "plots/optimal_stops/idw_final.png")

```


## Determining HDB Centroids

Now we have the fitered HDBs, we want to find out the centroids of the filtered HDB areas. This is so that we can then find out busstops which are most relevant to these HDBs. In order to accomplish this, we ran a KDE with a bandwidth of 1.4km on the filtered HDBs to view their clustering.


```{r kde start}

hdb_points <- as(hdb_yth_pts, "Spatial")
hdb_centers <- kde.points(hdb_points, h = 0.013) #1.4km bandwidth 
hdb_centers_sf <- st_as_sf(hdb_centers)

r_hdb_centers       <- raster(hdb_centers)
r_hdb_centers.m     <- mask(r_hdb_centers, base_spatial) 
  
kde_first <- tm_shape(r_hdb_centers.m) + 
  tm_raster(n=10, stretch.palette = FALSE,title="KDE Values") +
  tm_shape(base_map) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(hdb_yth_pts) + tm_dots(size=0.005) +
  tm_legend(legend.outside=TRUE) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="KDE of HDB locations",
            title.position = c('left', 'top'))

kde_first

tmap_save(kde_first, filename = "plots/optimal_stops/kde_first.png")

```


We then reclassified the kde values into 5 equal groups from 0 to 500 and removed the first 2 layers. This results in the following.

```{r kde second}

#Reclassify values in raster to make contour lines as seen in KDE plot.
reclass_values <- c(0,100,1, #reclassify kde values from 0-50 in group 1 and so on
                    100,200,2,
                    200,300,3,
                    300,400,4,
                    400,500,5)


reclass_hdb_centers <- reclassify(as(hdb_centers, "RasterLayer"), reclass_values) #reclassify kde values to groups
hdb_centers_poly <- rasterToPolygons(reclass_hdb_centers, dissolve = T, n = 16, digits = 3) #to make a polygon layer
hdb_centers_poly <- st_as_sf(hdb_centers_poly) #to make an SF object
hdb_centers_poly <- hdb_centers_poly[-c(1,2),] #remove polys with low kde values 
hdb_centers_poly <- st_cast(hdb_centers_poly,'POLYGON') #to split multipolygon to polygon to obtain centers

hdb_centers_poly$kde <- factor(hdb_centers_poly$kde) #to tell R to recognise the colours as categories

kde_second <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') + 
    tm_shape(hdb_yth_pts) + tm_dots(size = 0.01, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered KDEs with HDBs",
            title.position = c('left', 'top'))

kde_second

tmap_save(kde_second, filename = "plots/optimal_stops/kde_second.png")

```


Now it is hard to see the KDE polygons on that plot. To make the polygons easier to view, we plot the centroids of the individual layers to determine the central point of the HDB locations.


```{r KDE centroids}

#get the centroids of each polygon
hdb_centers_points <- st_centroid(hdb_centers_poly)

#HDB Centroids
kde_centroid <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') + 
    tm_shape(hdb_centers_points) + tm_dots(size = 0.1, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered KDEs with HDBs",
            title.position = c('left', 'top'))

kde_centroid

tmap_save(kde_centroid, filename = "plots/optimal_stops/kde_centroid.png")


```


## Bus Stop Mapping

The next set is to find out how the Bus Stops fare with our HDBs. First lets take a look at how the bus stops are distributed around Singapore.

```{r all bus}


bus_all <- tm_shape(base_map) + tm_borders() + 
  tm_shape(bus_spatial) + tm_bubbles(size = 0.15, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) + 
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Bus Stops in Singapore",
            title.size = 1,
            title.position = c('left', 'top'))

bus_all

tmap_save(bus_all, filename = "plots/optimal_stops/bus_all.png")

```


We overlay these with our KDEs earlier.

```{r all bus kde}

#Map to show how bustops fare with kdes
bus_all_kde <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') +
  tm_shape(bus_spatial) + tm_bubbles(size = 0.15, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) + 
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Bus Stops with KDEs",
            title.size = 1,
            title.position = c('left', 'top'))

bus_all_kde

tmap_save(bus_all_kde, filename = "plots/optimal_stops/bus_all_kde.png")

```


Next filter out only bus stops within these kdes. This step is not really necessary, but is doen to reduce computational intensity when calculating the best bus stop for each.

```{r filtered bus kde}

#remove busstops which are not inside the polygons
bus_inside_spatial <- st_intersection(bus_spatial, hdb_centers_poly)

#Show busstops which are inside the KDEs
bus_filtered_kde <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') +
  tm_shape(bus_inside_spatial) + tm_bubbles(size = 0.15, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) + 
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered Bus Stops with KDEs",
            title.size = 1,
            title.position = c('left', 'top'))

bus_filtered_kde

tmap_save(bus_filtered_kde, filename = "plots/optimal_stops/bus_filtered_kde.png")

```


Lastly, we find the closest bus stop to each centroid found earlier and assign each polygon a bus stop!

```{r filtered bus}


#find closest busstop to each centroid
closest_busstops_index <- st_nearest_feature(hdb_centers_points, bus_inside_spatial)
closest_busstops <- bus_inside_spatial[closest_busstops_index,]

#busstops 

bus_final <- tm_shape(base_map) + tm_borders() + 
  tm_shape(hdb_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') +
  tm_shape(closest_busstops) + tm_bubbles(size = 0.25, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) + 
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Final Bus Stops",
            title.size = 1,
            title.position = c('left', 'top'))

bus_final

tmap_save(bus_final, filename = "plots/optimal_stops/bus_final.png")

```


Now to help out with plotting optimal rouths, we want to tag each bus stop with an additional attribute stating what subzone it is in. We then can export this as a geojson file to be saved later.

```{r subzone tagging}

#Now we want to add the region attribute to each datapoint for each bus stop for later
#First extract out the polygons with the regions only
combined_base_map <- subzones %>%
  group_by(REGION_N) %>%
  summarise(geometry = st_combine(geometry)) 
combined_base_map <- st_transform(combined_base_map, crs = 4326)
combined_base_map <- st_make_valid(combined_base_map)

final_closest_busstops <- st_intersection(closest_busstops,combined_base_map) #get the regions for each busstop 
final_closest_busstops <- distinct(final_closest_busstops) #remove duplicated rows
final_closest_busstops <- final_closest_busstops %>% rename(Region = REGION_N)

#Export as a geojson such for bus line analysis
#st_write(final_closest_busstops, "bus_stops.geojson", driver = 'geoJSON',append = T) # DO NOT RUN AGAIN, IT WILL KEEP ADDING LAYERS 

```

We lastly want to evaluate how well our bus stops capture the HDBs in Singapore.

We do 2 forms of evaluation, one for all HDBS and one for our filtered HDBs.To then capture covered, we created a 1.5km buffer for each bus stop. The union of all busstops then determine our effectiveness in capturing the HDBs in Singapore. 

```{r bus stop eval}

#create buffers for each centroid
closest_busstops_buffer <- st_buffer(final_closest_busstops, dist = 1500) #1.5km buffer
closest_busstops_buffer <- st_union(closest_busstops_buffer) #join the buffers together
closest_busstops_buffer <- st_cast(closest_busstops_buffer,'POLYGON') #join the buffers together
closest_busstops_buffer <- st_make_valid(closest_busstops_buffer)


num_hdb_captured <- nrow(st_intersection(hdb_spatial, closest_busstops_buffer))
percentage_captured <- num_hdb_captured / nrow(hdb_spatial)
print(percentage_captured)
#1.5km radius: 47.1% HDB captured for all HDBs


num_hdb_captured <- nrow(st_intersection(hdb_yth_pts, closest_busstops_buffer))
percentage_captured <- num_hdb_captured / nrow(hdb_yth_pts)
print(percentage_captured)
#1.5km radius: 97.7% HDB captured for filtered HDBs



```

We can then visualise the evaluatation below.


```{r eval plots}


#Busstop buffers with ALL HDBs
eval_all <- tm_shape(base_map) + tm_borders() + 
  tm_shape(closest_busstops_buffer) + tm_polygons(col = 'yellow') + 
  tm_shape(hdb_spatial) + tm_bubbles(size = 0.05, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Bus Stop Buffers \nwith All HDBs",
            title.size = 1,
            title.position = c('left', 'top'))

eval_all

#Busstop buffers with Filtered HDBs
eval_filter <-tm_shape(base_map) + tm_borders() + 
  tm_shape(closest_busstops_buffer) + tm_polygons(col = 'yellow') + 
  tm_shape(hdb_yth_pts) + tm_bubbles(size = 0.05, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "bottom"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Bus Stop Buffers \nwith Filtered HDBs",
            title.size = 1,
            title.position = c('left', 'top'))

eval_filter


tmap_save(eval_all, filename = "plots/optimal_stops/eval_all.png")
tmap_save(eval_filter, filename = "plots/optimal_stops/eval_filter.png")

```


# Nightlife Bus Stop Analysis

We then repeated the same process with nightlife locations, except no IDW filtering was used here. A summarised version of the process is written below. 

A bandwidth of 750m was used and a buffer of 500m was used.

```{r nightlife}

#Onto finding ideal busstops for nightlife

#Central area of SG
central_area <- base_map %>% filter(PLN_AREA_N %in% c('DOWNTOWN CORE', 'SINGAPORE RIVER', 'MUSEUM','ROCHOR', 'KALLANG','OUTRAM')) %>% filter(!SUBZONE_N %in% c('TANJONG RHU', 'BENDEMEER','KALLANG BAHRU','KAMPONG BUGIS','BOON KENG','GEYLANG BAHRU'))
central_area <- st_make_valid(central_area)

nightlife_points <- as(nightlife, "Spatial")
nightlife_centers <- kde.points(nightlife_points, h = 0.00675) #750mm smoothing
nightlife_centers_sf <- st_as_sf(nightlife_centers)

r_nightlife_centers       <- raster(nightlife_centers)
r_nightlife_centers.m     <- mask(r_nightlife_centers, base_spatial) 
  
kde_n_first <- tm_shape(r_nightlife_centers.m) + 
  tm_raster(n=10, stretch.palette = FALSE,title="KDE Values") +
  tm_shape(base_map) + tm_borders() + tm_fill(alpha = 0) +
  tm_shape(nightlife) + tm_dots(size=0.2) +
  tm_legend(legend.outside=TRUE) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="KDE of nightlife locations",
            title.position = c('left', 'top'))

kde_n_first

tmap_save(kde_first, filename = "plots/optimal_stops/kde_n_first.png")


#reclassification for polygons
reclass_values_2 <- c(0,max(nightlife_centers$kde) / 6,1, #reclassify kde values from 0-1000 in group 1 and so on
                      max(nightlife_centers$kde) / 6,2 * max(nightlife_centers$kde) / 6,2,
                      2* max(nightlife_centers$kde) / 6,3 * max(nightlife_centers$kde) / 6,3,
                      3* max(nightlife_centers$kde) / 6,4 * max(nightlife_centers$kde) / 6,4,
                      4* max(nightlife_centers$kde) / 6,5 * max(nightlife_centers$kde) / 6,5,
                      5* max(nightlife_centers$kde) / 6,6 * max(nightlife_centers$kde) / 6,6)
                      #6* max(nightlife_centers$kde) / 7, max(nightlife_centers$kde),7)

reclass_nightlife_centers <- reclassify(as(nightlife_centers, "RasterLayer"), reclass_values_2) #reclassify kde values to groups
nightlife_centers_poly <- rasterToPolygons(reclass_nightlife_centers, dissolve = T, digits = 6) #to make a polygon layer
nightlife_centers_poly <- st_as_sf(nightlife_centers_poly) #to make an SF object
nightlife_centers_poly <- nightlife_centers_poly[-c(1,2),] #remove polys with low kde values 
nightlife_centers_poly <- st_cast(nightlife_centers_poly,'POLYGON') #to split multipolygon to polygon to obtain centers
nightlife_centers_poly$kde <- factor(nightlife_centers_poly$kde)


kde_n_second <- tm_shape(central_area) + tm_borders() + 
  tm_shape(nightlife_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') + 
    tm_shape(nightlife) + tm_dots(size = 0.2, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Filtered KDEs \nwith Nightlife",
            title.position = c('left', 'top'))

kde_n_second

tmap_save(kde_second, filename = "plots/optimal_stops/kde_second.png")

#get the centroids of each polygon
nightlife_centers_points <- st_centroid(nightlife_centers_poly)
#find closest busstop to each centroid
nightlife_centers_points <- st_transform(nightlife_centers_points, crs = 4326)
closest_nightlife_busstops_index <- st_nearest_feature(nightlife_centers_points, bus_spatial)
closest_nightlife_busstops <- bus_spatial[closest_nightlife_busstops_index,]

#bustops closest to each centroid
nightlife_final <- tm_shape(central_area) + tm_borders() + 
  tm_shape(nightlife_centers_poly) + tm_polygons(col = 'kde', title = 'KDE Group', palette = 'YlOrRd') +
  tm_shape(closest_nightlife_busstops) + tm_bubbles(size = 0.25, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) + 
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Final Bus Stops \n(Nightlife)",
            title.size = 1,
            title.position = c('left', 'top'))

nightlife_final

tmap_save(nightlife_final, filename = "plots/optimal_stops/nightlife_final.png")

#assigning location attribute to each busstop
#not needed as all are Central, but for completeness sake

combined_base_map_nightlife <- central_area %>%
  group_by(REGION_N) %>%
  summarise(geometry = st_combine(geometry)) 
combined_base_map_nightlife <- st_transform(combined_base_map_nightlife, crs = 4326)
combined_base_map_nightlife <- st_make_valid(combined_base_map_nightlife)

final_closest_busstops_nightlife <- st_intersection(closest_nightlife_busstops,combined_base_map_nightlife) #get the regions for each busstop 
final_closest_busstops_nightlife <- distinct(final_closest_busstops_nightlife) #remove duplicated rows
final_closest_busstops_nightlife <- final_closest_busstops_nightlife %>% rename(Region = REGION_N)

#st_write(final_closest_busstops_nightlife, "bus_stops_nightlife.geojson")

#Evaluation
#creating buffer
closest_nightlife_busstops_buffer <- st_buffer(final_closest_busstops_nightlife, dist = 500) #500m buffer
closest_nightlife_busstops_buffer <- st_union(closest_nightlife_busstops_buffer) #join the buffers together
closest_nightlife_busstops_buffer <- st_cast(closest_nightlife_busstops_buffer,'POLYGON') #join the buffers together
closest_nightlife_busstops_buffer <- st_make_valid(closest_nightlife_busstops_buffer)


#percentage covered
num_nightlife_captured <- nrow(st_intersection(nightlife, closest_nightlife_busstops_buffer))
percentage_captured <- num_nightlife_captured / nrow(nightlife)
print(percentage_captured)
#500m radius: 71.0% nightlife captured

#plotting the coverage

eval_nightlife <-tm_shape(central_area) + tm_borders() + 
  tm_shape(closest_nightlife_busstops_buffer) + tm_polygons(col = 'yellow') + 
  tm_shape(nightlife) + tm_bubbles(size = 0.5, scale = 0.5, col = 'black')  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "bottom"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Bus Stop Buffers \nwith Nightlife",
            title.size = 1,
            title.position = c('left', 'top'))

eval_nightlife

tmap_save(eval_nightlife, filename = "plots/optimal_stops/eval_nightlife.png")
```


The geojson files are then sent to the python code to determine optimal bus routes.

DONE! 




