---
title: "restaurants"
output: html_document
---

# 1. Loading all necessary libraries 
```{r setup}
library(GISTools)
library(raster)
library(tmap)
# library(rgdal) 
library(maptools) 
library(sp)
library(sf)
library(tidyverse)
library(rgeos)
library(ggplot2)
library(psych)
library(spatstat)
library(spdep)
library(jsonlite)
library(dplyr)
library(osmdata)
```

# 2. Helper functions
```{r helper functions}
# Create plot of basemap with specified regions, nightlife locations in sg and tmap mode
nightlife_basemap_plot <- function(base_plg, nightlife_sf,mode) {
  tmap_mode(mode)
  tmap_options(check.and.fix = TRUE)
  
  tm_shape(base_plg) + tm_borders(col='black') + 
    tm_shape(nightlife_sf) + tm_dots(size = 0.0075, col = "red") + # Nightlife locations
    tm_scale_bar(width = 0.15) +
    tm_layout(legend.format = list(digits = 0),
              legend.position = c("left", "bottom"),
              legend.text.size = 0.25, 
              legend.title.size = 0.5,
              title="Nightlife location in Singapore",
              title.position = c('left', 'bottom'))
}
```

# 3. Loading datasets
```{r data loading}
# 2.1 Basemap - Singapore Subzones (Polygons Vectors)
subzones <- st_read(dsn = "dataset/basemap", layer = "MP14_SUBZONE_NO_SEA_PL")
base_map <- st_read(dsn = "dataset/basemap", layer = "SGP_adm0")
base_map <- base_map %>% select(ISO)
base_map <- st_transform(base_map, crs=st_crs(subzones))

# 2.2 Nightlife (Bars + Clubs) Data (Points Vectors)
query <- opq(bbox = "singapore, singapore") %>% 
  add_osm_feature(key = "amenity", value = c("bar","biergarten", "pub", "nightclub"))
rawdata_nightlife <- osmdata_sf(query)
nightlife_points <- rawdata_nightlife$osm_points %>% drop_na(name) %>%
  dplyr::select(name, amenity, `addr:postcode`, `addr:street`) 

# some nightlife locations are represented as polygons, we convert them to point features
nightlife_plg <- rawdata_nightlife$osm_polygons %>% drop_na(name) %>%
  dplyr::select(name, amenity, `addr:postcode`, `addr:street`) %>%
  st_centroid()
nightlife <- rbind(nightlife_points, nightlife_plg)

# Create a Nightlife Spatial Points DataFrame
nightlife_sf <- st_transform(nightlife, crs=st_crs(subzones)) # project to basemap's CRS
nightlife_sf <- st_intersection(nightlife_sf, base_map) # remove nightlife points that lie outside of basemap

nightlife_basemap_plot(base_map, nightlife_sf, "view")
```

# 4. Filter regions of interest (central regions of Singapore)
We are choosing to narrow our scope to nightlife locations in the central regions of Singapore. After close inspection of the basemap polygons, we define these central regions as Downtown Core, Singapore River, Rochor, Outram and Lavendar.
```{r , echo=TRUE}
# Filter polygons of interest
central_plg <- subzones %>% filter(PLN_AREA_C %in% c("DT", "SR", "RC", "OT") | SUBZONE_N %in% c("LAVENDER"))

# Plot with filtered polygons
nightlife_basemap_plot(central_plg, nightlife_sf, "view")
View(subzones)

```

# 4. Filter nightlife locations that lie in filtered polygons
```{r , echo=TRUE}
nightlife_filtered_sf <- st_intersection(nightlife_sf, central_plg)

nightlife_basemap_plot(base_map, nightlife_filtered_sf, "view")
```

# 4. Save nightlife and basemap
```{r}
st_write(nightlife_filtered_sf, "dataset/processed_datasets/nightlife_locations", driver = "ESRI Shapefile")
```