---
title: "Story 1: Supper"
output:
  pdf_document: default
  html_document: default
date: "2023-10-09"
---
## 1. Setting up
```{r setup, include=FALSE}
library(GISTools)
library(raster)
library(tmap)
library(rgdal) 
library(maptools) 
library(sp)
library(sf)
library(tidyverse)
library(rgeos)
library(ggplot2)
library(psych)
library(spatstat)
library(spdep)
library(jsonlite)
library(dplyr)
library(gstat)

```

```{r basemap}
subzones <- st_read(dsn = "dataset/basemap", layer = "MP14_SUBZONE_NO_SEA_PL")
#Central area of SG
base_map <- st_read(dsn = "dataset/basemap", layer = "MP14_SUBZONE_NO_SEA_PL")
central_area <- base_map %>% filter(PLN_AREA_N %in% c('DOWNTOWN CORE', 'BUKIT MERAH', 'SINGAPORE RIVER', 'MUSEUM', 'RIVER VALLEY', 'ORCHARD','NEWTON','ROCHOR','TANGLIN', 'KALLANG','OUTRAM')) %>% filter(!SUBZONE_N %in% c( 'CITY TERMINALS','MARITIME SQUARE','ALEXANDRA HILL','TELOK BLANGAH DRIVE','DEPOT ROAD','TELOK BLANGAH WAY','TELOK BLANGAH RISE'))
base_map <- st_transform(central_area, crs=st_crs(central_area)) 
```

### 2. Distribution of Nightlife locations and latenight Restaurants

```{r Nightlife-Restaurants Central distribution}
# Load filtered Nightlife - Restaurants Data

# Load nightlife locations
nightlife_sf <- st_read("dataset/processed_datasets/nightlife_locations.geojson", crs=st_crs(subzones))
nightlife_sf <- st_transform(nightlife_sf, crs=st_crs(subzones)) # project to basemap's CRS

# Load restaurants
restaurants_sf <- st_read("dataset/processed_datasets/restaurants.geojson", crs=st_crs(subzones))
restaurants_sf <- st_transform(restaurants_sf, crs=st_crs(subzones)) # project to basemap's CRS

nightlife_filtered_sf = nightlife_sf
nightlife_filtered_sf$label = 'Nightlife locations'
nightlife_filtered_sf <- nightlife_filtered_sf %>% select(label)

midnight_restaurant_filtered_sf = restaurants_sf %>% filter(midnight_restaurant=="True")
midnight_restaurant_filtered_sf$label = 'Late night restaurants'
midnight_restaurant_filtered_sf <- midnight_restaurant_filtered_sf %>% select(label)

combined <- rbind(nightlife_filtered_sf, midnight_restaurant_filtered_sf)

tm_shape(central_area) + tm_polygons(col="white") + 
  tm_shape(combined) + tm_dots(size= 0.15, col = "label", palette = c("green2", "blue")) + 
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Nightlife and Latenight restaurants at Central Area",
            title.position = c('left', 'top'))
```

### 3. Hypothesis Testing of 2nd order effect

``` {r PPA}
  
# Point pattern analysis
nightlife_ppp <- as.ppp(nightlife_filtered_sf)
marks(nightlife_ppp) <- "nightlife"
nightlife_ppp$n

latenight_restaurants_ppp <- as.ppp(midnight_restaurant_filtered_sf)
marks(latenight_restaurants_ppp) <- "latenight restaurants"
latenight_restaurants_ppp$n


P <- superimpose(nightlife_ppp, latenight_restaurants_ppp)

# hypothesis testing for the distribution of nightlife and latenight restaurants
nn.sim = vector()
P.r = P
for(i in 1:999){
  marks(P.r) = sample(P$marks)  # Reassign labels at random, point locations don't change
  nn.sim[i] = mean(nncross(P.r[P.r$marks == "nightlife", ] ,P.r[P.r$marks == "latenight restaurants", ])$dist)
}
hist(nn.sim,breaks=30, xlim=c(50,200), main="Histogram of nearest neighbour distance")
actual_dist = mean(nncross(nightlife_ppp,latenight_restaurants_ppp)$dist)
abline(v=actual_dist,col="red")
axis(1, at=actual_dist,labels=round(actual_dist,1))

```

As the p-value is 1, there is strong evidence that the observed distribution of nightlife locations and late-night restaurants is unlikely to have occurred by random chance.

### 4.Crowd Density Interpolation

```{r bcr data}

#Bars and Clubs Crowd Density dataframe
bcr <- read_csv('dataset/besttime/processed_bars_clubs.csv')
bcr <- na.omit(bcr)
#Restaurants Crowd Density dataframe
res <- read_csv('dataset/besttime/processed_restaurants.csv')
res <- na.omit(res)
res <- res %>% rename(`12am_cd` = `12pm_cd`)

bcr <- bcr %>% select(lat,lon, `10pm_cd`,`11pm_cd`,`12am_cd`,`1am_cd`,`2am_cd`) %>% mutate(cat = "Bars and Clubs")
res <- res %>% select(lat,lon, `10pm_cd`,`11pm_cd`,`12am_cd`,`1am_cd`,`2am_cd`)%>% mutate(cat = "Restaurants")
bcr <- rbind(bcr,res)

bcr$`cd_10` <- bcr$`10pm_cd`
bcr$`cd_11` <- bcr$`11pm_cd`
bcr$`cd_12` <- bcr$`12am_cd`
bcr$`cd_1` <- bcr$`1am_cd`
bcr$`cd_2` <- bcr$`2am_cd`

coordinates(bcr) <- ~lon + lat


#Tranform to spatial dataframe and see how it looks like
plot(bcr) #notice that one point is so much further away than the rest
bcr <- bcr[-c(10,49),] #Remove that one point for better interpolation
plot(bcr) #Corrected plot without annoying point

bcr$X <- coordinates(bcr)[,1]
bcr$Y <- coordinates(bcr)[,2]

bcr <- bcr[-zerodist(bcr)[,1],]

base_spatial <- as(base_map, 'Spatial')


bcr_spatial <- st_as_sf(bcr, coords = c("lon", "lat"))
```


```{r grid}

#Making the raster grid

grd              <- as.data.frame(spsample(bcr, "regular", n=50000))
names(grd)       <- c("X", "Y")
coordinates(grd) <- c("X", "Y")
gridded(grd)     <- TRUE
fullgrid(grd)    <- TRUE

crs(grd) <- crs(bcr)

```


```{r interpolates}

#10pm
f.bcr <- as.formula(cd_10 ~ lon + lat) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.0025)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.025, nugget=600))

# The following plot allows us to assess the fit
vari_plot_10 <- plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.05))
vari_plot_10


f.bcr <- as.formula(cd_10 ~ X + Y) 

dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)

#tm_shape(base_map) + tm_borders(col='black') + tm_fill(col = "white") +
i_10 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Interpolation of Crowd Density at 10pm", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120),alpha = 0.7) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)


#11pm

f.bcr <- as.formula(cd_11 ~ lon + lat) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.0025)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.025, nugget=650))

# The following plot allows us to assess the fit
vari_plot_11 <- plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.05))
vari_plot_11


f.bcr <- as.formula(cd_11 ~ X + Y) 

dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)

#tm_shape(base_map) + tm_borders(col='black') + tm_fill(col = "white") +
i_11 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Interpolation of Crowd Density at 11pm", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120),alpha = 0.7) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)

#12am

f.bcr <- as.formula(cd_12 ~ lon + lat) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.0025)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.025, nugget=650))

# The following plot allows us to assess the fit
vari_plot_12 <- plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.05))
vari_plot_12


f.bcr <- as.formula(cd_12 ~ X + Y) 

dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)

#tm_shape(base_map) + tm_borders(col='black') + tm_fill(col = "white") +
i_12 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Interpolation of Crowd Density at 12am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120),alpha = 0.7) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)

#1am

f.bcr <- as.formula(cd_1 ~ lon + lat) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.0025)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.025, nugget=650))

# The following plot allows us to assess the fit
vari_plot_1 <- plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.05))
vari_plot_1


f.bcr <- as.formula(cd_1 ~ X + Y) 

dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)

#tm_shape(base_map) + tm_borders(col='black') + tm_fill(col = "white") +
i_1 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Interpolation of Crowd Density at 1am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120),alpha = 0.7) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)

#2am

f.bcr <- as.formula(cd_2 ~ lon + lat) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.0025)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.025, nugget=650))

# The following plot allows us to assess the fit
vari_plot_2 <- plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.05))
vari_plot_2


f.bcr <- as.formula(cd_2 ~ X + Y) 

dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)

#tm_shape(base_map) + tm_borders(col='black') + tm_fill(col = "white") +
i_2 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Interpolation of Crowd Density at 1am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120),alpha = 0.7) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)





```


```{r 2bcr}

f.bcr <- as.formula(cd_2 ~ 1) 

var.smpl.bcr <- variogram(f.bcr, bcr, cloud = FALSE, cutoff=1, width=0.001)

dat.fit.bcr  <- fit.variogram(var.smpl.bcr, fit.ranges = F, fit.sills = F,
                          vgm(psill=400, model="Wav", range=0.015, nugget=300))

# The following plot allows us to assess the fit
plot(var.smpl.bcr, dat.fit.bcr, xlim = c(0, 0.06))



f.bcr <- as.formula(cd_2 ~ 1) 


dat.krg <- krige(f.bcr, bcr, grd, dat.fit.bcr)

r <- raster(dat.krg)
# r.m <- mask(r, W)

tm_shape(r) + 
  tm_raster(n=10, stretch.palette=FALSE, 
            title="Predicted Crowd Density at 2am", palette = 'RdBu', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120), midpoint = 0) +
  tm_shape(bcr) + tm_dots(size=0.05, col = 'cd_2', palette = 'RdBu', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120), midpoint = 0) +
  tm_legend(legend.outside=TRUE)

```




```{r trends}

#install.packages('magick')
library(magick)
f.bcr <- as.formula(cd_10 ~ X + Y) 

lm.1 <- lm(f.bcr, data = bcr)

dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 
r   <- raster(dat.1st)

l_1 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Linear Trend of Crowd Density at 10pm", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120)) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)


f.bcr <- as.formula(cd_11 ~ X + Y) 

lm.1 <- lm(f.bcr, data = bcr)

dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 
r   <- raster(dat.1st)

l_2 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Linear Trend of Crowd Density at 11pm", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120)) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)


f.bcr <- as.formula(cd_12 ~ X + Y) 

lm.1 <- lm(f.bcr, data = bcr)

dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 
r   <- raster(dat.1st)

l_3 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Linear Trend of Crowd Density at 12am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120)) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)

f.bcr <- as.formula(cd_1 ~ X + Y) 

lm.1 <- lm(f.bcr, data = bcr)

dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 
r   <- raster(dat.1st)

l_4 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Linear Trend of Crowd Density at 1am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120)) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)

f.bcr <- as.formula(cd_2 ~ X + Y) 

lm.1 <- lm(f.bcr, data = bcr)

dat.1st <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata=grd))) 
r   <- raster(dat.1st)

l_5 <- tm_shape(r) + 
  tm_raster(n=13, stretch.palette=FALSE, 
            title="Linear Trend of Crowd Density at 2am", palette = 'Reds', style = 'fixed',breaks = c(-10,0,10,20,30,40,50,60,70,80,90,100,110,120)) +
  tm_shape(bcr) + tm_dots(size=0.1, col = 'cat', palette = c('blue','green2')) +
  tm_legend(legend.outside=TRUE)


```


```{r trend animations}


tmap_save(l_1, filename = "plots/crowd_density/linear trend at 10pm.png")
tmap_save(l_2, filename = "plots/crowd_density/linear trend at 11pm.png")
tmap_save(l_3, filename = "plots/crowd_density/linear trend at 12am.png")
tmap_save(l_4, filename = "plots/crowd_density/linear trend at 1am.png")
tmap_save(l_5, filename = "plots/crowd_density/linear trend at 2am.png")

img_list <- image_read(c("plots/crowd_density/linear trend at 10pm.png","plots/crowd_density/linear trend at 11pm.png","plots/crowd_density/linear trend at 12am.png","plots/crowd_density/linear trend at 1am.png","plots/crowd_density/linear trend at 2am.png"))
animation <- image_animate(img_list, fps = 1)  # Adjust fps as needed
image_write(animation, path = "plots/crowd_density/linear trends.gif")


```



## 5. KDE to recommend new late night restaurants

```{r kde-of-nightlife}
nightlife_points <- as(nightlife_sf, "Spatial")
nightlife_centers <- kde.points(nightlife_points, h = 500)
nightlife_centers_sf <- st_as_sf(nightlife_centers)

tmap_mode("plot")
tm_shape(nightlife_centers) + tm_raster()
```


```{r get-kde-polygons}
reclass_values <- c(0,0.0000002,1, #reclassify kde values from 0-0.0000002 in group 1 and so on
                    0.0000002,0.0000004,2,
                    0.0000004,0.0000006,3,
                    0.0000006,0.0000008,4,
                    0.0000008,0.0000010,5,
                    0.0000010,0.0000012,6)

reclass_nightlife_centers <- reclassify(as(nightlife_centers, "RasterLayer"), reclass_values) #reclassify kde values to groups
nightlife_centers_poly <- rasterToPolygons(reclass_nightlife_centers, dissolve = T) #to make a polygon layer
nightlife_centers_poly <- st_as_sf(nightlife_centers_poly) #to make an SF object
nightlife_centers_poly <- nightlife_centers_poly[-c(1),] #remove polys with low kde values 
nightlife_centers_poly <- st_cast(nightlife_centers_poly,'POLYGON') #to split multipolygon to polygon to obtain centers
```

```{r plot-kde}
nightlife_centers_poly <- st_transform(nightlife_centers_poly, crs=st_crs(subzones)) # project to basemap's CRS

# Plot of nightlife location and density polygons
tm_shape(base_map) + tm_borders() + 
  tm_shape(nightlife_centers_poly) + tm_fill(col = 'kde') + 
  tm_shape(nightlife_sf) + tm_bubbles(size = 0.001, col = "black")  +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="KDE of Nightlife locations in Central Singapore",
            title.position = c('left', 'top'))
```

```{r get restaurants-in-kde-region}
# Filter for nightlife locations within KDE regions
all_kde <- st_union(nightlife_centers_poly)
nightlife_kde <- st_intersection(all_kde, nightlife_sf)
print(length(nightlife_kde))
print(nrow(nightlife_sf))
print(sprintf("A total of %d nightlife locations lie in the filtered KDE regions, out of %d original nightlife locations. That means %f of nightlife locations lie in the filtered KDE regions.", length(nightlife_kde), nrow(nightlife_sf), length(nightlife_kde)/nrow(nightlife_sf)))
```
```{r recommended-restaurants}

midnight_restaurants <- restaurants_sf %>%
  filter(midnight_restaurant == "True")

non_midnight_restaurants <- restaurants_sf %>%
  filter(midnight_restaurant == "False")

# Combine polygons by KDE class and find its intersection with restaurants
grouped_union <- nightlife_centers_poly %>%
  group_by(kde) %>%
  summarise(geometry = st_union(geometry))

nightlife_intersection <- st_intersection(grouped_union, nightlife_sf)

kde_nightlife_count <- nightlife_intersection %>%
  group_by(kde) %>%
  summarise(nightlife_count = n())

print(kde_nightlife_count)

# Combine polygons by KDE class and find its intersection with restaurants
restaurant_intersection <- st_intersection(grouped_union, non_midnight_restaurants)
midnight_restaurant_intersection <- st_intersection(grouped_union, midnight_restaurants)

kde_restaurant_count <- restaurant_intersection %>%
  group_by(kde) %>%
  summarise(restaurant_count = n())

# View the result
print(kde_restaurant_count)
```

```{r}
# Plot of all non-midnight restaurants within the extracted polygons
tm_shape(restaurant_intersection) +
  tm_bubbles(size = 0.1, col = "kde") +
  tm_shape(midnight_restaurant_intersection) +
  tm_bubbles(size = 0.1, col = "black") +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_layout(legend.format = list(digits = 0),
            legend.position = c("left", "top"),
            legend.text.size = 0.5, 
            legend.title.size = 1,
            title="Recommended Restaurants in Central Singapore",
            title.position = c('left', 'top'))
```